\section{Sets and Counting Puzzles}

Find the indicated sets.
\begin{puzzle}
    List all subsets of the following set.
    \[\{Al, Bob\}\]
\end{puzzle}

\begin{puzzle}
    List all subsets of the following set.
    \[\{Al, Bob, Chris\}\]
\end{puzzle}

\begin{puzzle}
    List the elements of the following set.
    \[\{Al, Bob, Chris, Dave\} \cap \{Bob, Chris, Dave, Ed\}\]
\end{puzzle}

\begin{puzzle}
    List the elements of the following set.
    \[\{Al, Bob, Chris, Dave\} \cup \{Bob, Chris, Dave, Ed\}\]
\end{puzzle}

In Problems \ref{puzzle_first_VW_set} - \ref{puzzle_last_VW_set}, let Universal set \( U = \{a, b, c, d, e, f, g, h, i, j\} \), \( V = \{a, e, i, f, h\} \), and \( W = \{a, c, e, g, i\} \). List the members of the following sets.

\begin{puzzle}\label{puzzle_first_VW_set}

    \[V \cup W\]
\end{puzzle}

\begin{puzzle}
    \[V \cap W\]
\end{puzzle}

\begin{puzzle}
    \[\overline{V \cup W}\]
\end{puzzle}

\begin{puzzle}\label{puzzle_last_VW_set}
    \[\overline{V} \cap \overline{W}\]
\end{puzzle}

In problems \ref{puzzle_first_AB_set} - \ref{puzzle_last_AB_set}, let Universal set \( U = \{1,2,3,4,5,6,7,8,9,10\} \), \( A = \{1,2,3,4,5\} \), \( B = \{1,3,4,6\} \), and \( C = \{2,4,6\} \).

List the members of the following sets.


\begin{puzzle}\label{puzzle_first_AB_set}
    \[A \cup B\]
\end{puzzle}

\begin{puzzle}
    \[A \cap C\]
\end{puzzle}

\begin{puzzle}
    \[\overline{A \cup B} \cap C\]
\end{puzzle}

\begin{puzzle}\label{puzzle_last_AB_set}
    \[\overline{A} \cup \overline{B \cap C}\]
\end{puzzle}

Find the number of elements in the following sets.

\begin{puzzle}
    In Mrs. Yamamoto's class of 35 students, 12 students are taking history, 18 are taking English, and 4 are taking both. Draw a Venn diagram and determine how many students are taking neither history nor English?
\end{puzzle}

\begin{puzzle}
    In the County of Santa Clara 700,000 people read the San Jose Mercury News, 400,000 people read the San Francisco Examiner, and 100,000 read both newspapers. How many read either the Mercury News or the Examiner?
\end{puzzle}

\begin{puzzle}
    A survey of athletes revealed that for their minor aches and pains, 30 used aspirin, 50 used ibuprofen, and 15 used both. How many athletes were surveyed?
\end{puzzle}