\section{Chapter Review}

\begin{puzzle}
    Find an equation of the x-axis.
\end{puzzle}

\begin{puzzle}
    Find the slope of the line whose equation is \(2x + 3y = 6\).

\end{puzzle}

\begin{puzzle}
    Find the slope of the line whose equation is \(y = -3x + 5\).
\end{puzzle}

\begin{puzzle}
    Find both the x and y intercepts of the line \(3x - 2y = 12\).
\end{puzzle}

\begin{puzzle}
    Find an equation of the line whose slope is 3 and y-intercept 5.
\end{puzzle}

\begin{puzzle}
    Find an equation of the line whose x-intercept is 2 and y-intercept 3.
\end{puzzle}

\begin{puzzle}
    Find an equation of the line that has slope 3 and passes through the point \((2, 15)\).
\end{puzzle}

\begin{puzzle}
    Find an equation of the line that has slope \(-\frac{3}{2}\) and passes through the point \((4, 3)\).
\end{puzzle}

\begin{puzzle}
    Find an equation of the line that passes through the points (0, 32) and (100, 21).
\end{puzzle}

\begin{puzzle}
    Find an equation of the line that passes through the point (2, 5) and is parallel to the line \( y = 3x+4 \).
\end{puzzle}

\begin{puzzle}
    Find the point of intersection of the lines \( 2x - 3y = 9 \) and \( 3x + 4y = 5 \).
\end{puzzle}

\begin{puzzle}
    Is the point (3, -2) on the line \( 5x - 2y = 11 \)?
\end{puzzle}

\begin{puzzle}
    Find two points on the line given by the parametric equations, \( x = 2 + 3t, y = 1 - 2t \).
\end{puzzle}

\begin{puzzle}
    Find two points on the line \( 2x - 6 = 0 \).
\end{puzzle}

\begin{puzzle}
    Graph the line $2x - 3y + 6 = 0$.
\end{puzzle}

\begin{puzzle}
    Graph the line $y = -2x + 3$.
\end{puzzle}

\begin{puzzle}
    A  college student who is 60 inches tall weighs 100 pounds, and another student who is 66 inches tall weighs 124 pounds. Assuming the relationship between the students’ weights and heights is linear, write an equation giving the relationship between heights and weights of students, and use this relationship to predict the weight of a student who is 70 inches tall.
\end{puzzle}

\begin{puzzle}
    In deep-sea diving, the pressure exerted by water plays a great role in designing underwater equipment. If at a depth of 10 feet there is a pressure of 21 lb/in$^2$, and at a depth of 50 ft there is a pressure of 75 lb/in$^2$, write an equation giving a relationship between depth and pressure. Use this relationship to predict pressure at a depth of 100 ft.
\end{puzzle}

\begin{puzzle}
    If the variable cost to manufacture an item is \$30, and the fixed costs are \$2750, write the cost function.
\end{puzzle}

\begin{puzzle}
    The variable cost to manufacture an item is \$10, and it costs \$2,500 to produce 100 items. Write the cost function, and use this function to estimate the cost of manufacturing 300 items.
\end{puzzle}

\begin{puzzle}
    If it costs \$2,100 to manufacture 100 items of a product, and \$4,200 to manufacture 200 items. If $x$ represents the number of items, and $y$ the costs, find the cost equation, and use this function to predict the cost of 1,000 items.
\end{puzzle}

\begin{puzzle}
    In 1980, the average house in Palo Alto cost \$280,000, and in 1997 costs \$450,000. Assuming a linear relationship, write an equation that will give the price of the house in any year, and use this equation to predict the price of a similar house in the year 2010.
\end{puzzle}

\begin{puzzle}
    The population of Argentina in 1987 was 31.5 million and in 1997 it was 42.5 million. Assuming a linear relationship, write an equation that will give the population of Argentina in any year, and use this equation to predict the population of Argentina in the year 2010.
\end{puzzle}

\begin{puzzle}
    In 1955, an average new Chevrolet sold for \$2,400, and a similar Chevrolet sold for \$15,000 in 1995. Assuming a linear relationship, write an equation that will give the price of an average Chevrolet in any year. Use this equation to predict the price of an average Chevrolet in the year 2010.
\end{puzzle}

\begin{puzzle}
    Two hundred items are demanded at a price of \$5, and 300 items are demanded at a price of \$3. If \( x \) represents the price, and \( y \) the number of items, write the demand function.
\end{puzzle}

\begin{puzzle}
    A supply curve for a product is the number of items of the product that can be made available at different prices. A manufacturer of Tickle Me Elmo dolls can supply 2000 dolls if the dolls are sold for \$30 each, but he can supply only 800 dolls if the dolls are sold for \$10 each. If \( x \) represents the price of dolls and \( y \) the number of items, write an equation for the supply curve.
\end{puzzle}

\begin{puzzle}
    Suppose you are trying to decide on a price for your latest creation - a coffee mug that never tips. Through a survey, you have determined that at a price of \$2, you can sell 2100 mugs, but at a price of \$12 you can only sell 100 mugs. Furthermore, your supplier can supply you 3500 mugs if you charge your customers \$12, but only 500 if you charge \$2. What price should you charge so that the supply equals demand, and at that price how many coffee mugs will you be able to sell?
\end{puzzle}

\begin{puzzle}
    A car rental company offers two plans. Plan I charges \$12 a day and 12 cents a mile, while Plan II charges \$30 a day but no charge for miles. If you were to drive 300 miles in a day, which plan is better? For what mileage are both rates the same?
\end{puzzle}

\begin{puzzle}
    The supply curve for a product is \( y = 25x - 1,000 \) and the demand curve for the same product is \( y = -35x + 8,000 \), where \( x \) is the price and \( y \) the number of items produced. Find the following:
    \begin{enumerate}
        \item At a price of \$10, how many items will be in demand?
        \item At what price will 4,000 items be supplied?
        \item What is the equilibrium price for this product?
        \item How many items will be manufactured at the equilibrium price?
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    The supply curve for a product is \( y = 625x - 600 \) and the demand curve for the same product is \( y = -125x + 8,400 \), where \( x \) is the price and \( y \) the number of items produced. Find the equilibrium price and determine the number of items that will be produced at that price.
\end{puzzle}

\begin{puzzle}
    Both Jenny and Masur are sales people for Athletic Shoes. Jenny gets paid \$8 per hour plus 4\% commission on the sales. Masur gets paid \$10 per hour plus 8\% commission on the sales in excess of \$1,000. If they work 8-hour days, for what sales amount would they both earn the same daily amounts?
\end{puzzle}

\begin{puzzle}
    A company’s revenue and cost in dollars are given by \( R(x) = 25x + 9,000 \) and \( C(x) = 10x + 9,000 \), where \( x \) represents the number of items. Find the number of items that must be produced to break even.
\end{puzzle}

\begin{puzzle}
    A firm producing video tapes has fixed costs of \$6,800, and a variable cost of 30 cents per tape. If the video tapes sell for \$2 each, find the number of tapes that must be produced to break-even.
\end{puzzle}

\begin{puzzle}
    A firm producing disposable cameras has fixed costs of \$8,000, and variable cost of 50 cents a camera. If the cameras sell for \$3.50, how many cameras must be produced to break-even?
\end{puzzle}

\begin{puzzle}
    The Stanley Company is coming up with a new cordless travel shaver just before the Christmas holidays. It hopes to sell 10,000 of these shavers in the month of December alone. The manufacturing variable cost is \$3 and the fixed costs \$100,000. If the shavers sell for \$11 each, how many must be produced to break-even?
\end{puzzle}
