\section{Classification of Finance Puzzles}

Let the letters \( A, B, C, D, E \) and \( F \) be represented as follows:
\begin{align*}
    A & = \text{FV of a lump-sum}     \\
    B & = \text{PV of a lump-sum}     \\
    C & = \text{FV of an annuity}     \\
    D & = \text{Sinking fund payment} \\
    E & = \text{Installment payment}  \\
    F & = \text{PV of an annuity}
\end{align*}
For the following problems, Classify, by writing the appropriate letter in the box, and write an equation for solution.

\begin{puzzle}
    What monthly deposits made to an account paying 9\% will grow to \$10,000 in 4 years?
\end{puzzle}

\begin{puzzle}
    An amount of \$4000 is invested at 6\% compounded daily. What will the final amount be in 5 years?
\end{puzzle}

\begin{puzzle}
    David has won a lottery paying him \$10,000 per month for the next 20 years. He'd rather have the whole amount in one lump sum now. If the current interest rate is 7\%, how much money can he hope to get?
\end{puzzle}

\begin{puzzle}
    Each month Linda deposits \$250 in an account that pays 9\%. How much money will she have in 4 years?
\end{puzzle}

\begin{puzzle}
    Find the monthly payment for a \$15,000 car if the loan is amortized over 4 years at a rate of 10\%.
\end{puzzle}

\begin{puzzle}
    What lump-sum deposited in an account paying 7\% compounded daily will grow to \$10,000 in 5 years?
\end{puzzle}

\begin{puzzle}
    What amount of quarterly payments will amount to \$250,000 in 5 years at a rate of 8\%?
\end{puzzle}

\begin{puzzle}
    The Chang family bought their house 25 years ago. They had their loan financed for 30 years at an interest rate of 11\% resulting in a payment of \$1350 a month. Find the balance of the loan.
\end{puzzle}

\begin{puzzle}
    A 10-year \$1000 bond pays \$35 every six months. If the current interest rate is 8\%, in order to find the fair market value of the bond, we need to find the following:
    \begin{enumerate}
        \item The present value of \$1000.
        \item The present value of the \$35 per six month payments.
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    What lump-sum deposit made today is equal to 33 monthly deposits of \$500 if the interest rate is 8\%?
\end{puzzle}

\begin{puzzle}
    What should be the size of the monthly deposits made to an account paying 10\% so that their accumulated value will be \$10,000 in six years?
\end{puzzle}

\begin{puzzle}
    A department store charges a finance charge of 1.5\% per month on the outstanding balance. If Ned charged \$400 three months ago and has not paid his bill, how much does he owe?
\end{puzzle}

\begin{puzzle}
    What will the value of \$300 monthly deposits be in 10 years if the account pays 12\% compounded monthly?
\end{puzzle}

\begin{puzzle}
    What lump-sum deposited at 6\% compounded daily will grow to \$2000 in three years?
\end{puzzle}

\begin{puzzle}
    A company buys an apartment complex for \$5,000,000 and amortizes the loan over 10 years. What is the yearly payment if the interest rate is 14\%?
\end{puzzle}

\begin{puzzle}{10.85}
    In 1970, a house in Reno cost \$23,000. If the inflation rate is 8\%, what is the price of that house in 1997?
    \end{puzzle}
    
    \begin{puzzle}{10.86}
    You determine that you can afford to pay \$400 per month for a car. What is the maximum price you can pay for a car if the interest rate is 11\% and you want to repay the loan in 4 years?
    \end{puzzle}
    
    \begin{puzzle}{10.87}
    A business needs \$350,000 in 5 years. How much lump-sum should be put aside in an account that pays 9\% so that five years from now the company will have \$350,000?
    \end{puzzle}
    
    \begin{puzzle}{10.88}
    A person wishes to have \$500,000 in a pension fund 20 years from now. How much should he deposit each month in an account paying 9\% compounded monthly?
    \end{puzzle}