\section{More Applications}

Solve the following puzzles.

\begin{puzzle}
    Solve for \( x \) and \( y \).
    \[ y = 3x + 4 \]
    \[ y = 5x - 2 \]

\end{puzzle}

\begin{puzzle}
    Solve for \( x \) and \( y \).
    \[ 2x - 3y = 4 \]
    \[ 3x - 4y = 5 \]

\end{puzzle}

\begin{puzzle}
    The supply curve for a product is \( y = 2000x + 13000 \), and the demand curve is \( y = -1000x + 28000 \), where \( x \) represents the price and \( y \) the number of items. At what price will the supply equal demand, and how many items will be produced at that price?

\end{puzzle}

\begin{puzzle}
    The supply curve for a product is \( y = 300x + 9000 \), and the demand curve is \( y = -100x + 14000 \), where \( x \) represents the price and \( y \) the number of items. At what price will the supply equal demand, and how many items will be produced at that price?

\end{puzzle}

\begin{puzzle}
    A demand curve for a commodity is the number of items the consumer will buy at different prices. It has been determined that at a price of \( \$2 \) a store can sell \( 2400 \) of a particular type of toy dolls, and for a price of \( \$8 \) the store can sell \( 600 \) such dolls. If \( x \) represents the price of dolls and \( y \) the number of items sold, write an equation for the demand curve.
\end{puzzle}


\begin{puzzle}
    A supply curve for a commodity is the number of items of the product that can be made available at different prices. A manufacturer of toy dolls can supply 2000 dolls if the dolls are sold for \$8 each, but he can supply only 800 dolls if the dolls are sold for \$2 each. If \( x \) represents the price of dolls and \( y \) the number of items, write an equation for the supply curve.

\end{puzzle}

\begin{puzzle}
    The equilibrium price is the price where the supply equals the demand. From the demand and supply curves obtained in the previous two problems, find the equilibrium price, and determine the number of items that can be sold at that price.

\end{puzzle}

\begin{puzzle}
    A car rental company offers two plans. Plan I charges \$10 a day and 10 cents a mile, while Plan II charges 14 cents a mile, but no flat fee. If you were to drive 300 miles in a day, which plan is better? For what mileage are both rates equal?
\end{puzzle}


\begin{puzzle} A break-even point is the intersection of the cost function and the revenue function, that is, where the total cost equals revenue. Mrs. Jones Cookies Store’s revenue and cost in dollars for \( x \) number of cookies is given by \( R = 0.80x \) and \( C = 0.05x + 3000 \). Find the number of cookies that must be sold so that the revenue and cost are the same.

\end{puzzle}

\begin{puzzle} A company’s revenue and cost in dollars are given by \( R = 225x \) and \( C = 75x + 6000 \), where \( x \) is the number of items. Find the number of items that must be produced to break-even.

\end{puzzle}

\begin{puzzle} A firm producing computer diskettes has a fixed costs of \$10,725, and variable cost of 20 cents a diskette. Find the break-even point if the diskettes sell for \$1.50 each.

\end{puzzle}

\begin{puzzle} Whackemhard Sports is planning to introduce a new line of tennis rackets. The fixed costs for the new line are \$25,000 and the variable cost of producing each racket is \$60. If the racket sells for \$80, find the number of rackets that must be sold in order to break even.
\end{puzzle}