\section{Combination Puzzles with Several Sets}

Following problems involve combinations from several different sets

\begin{puzzle}
    How many 5-people committees consisting of three boys and two girls can be chosen from a group of four boys and four girls?
\end{puzzle}

\begin{puzzle}
    A club has 4 men, 5 women, 8 boys and 10 girls as members. In how many ways can a group of 2 men, 3 women, 4 boys and 4 girls be chosen?
\end{puzzle}

\begin{puzzle}
    How many 4-people committees chosen from four men and six women will have at least three men?
\end{puzzle}

\begin{puzzle}
    A batch contains 10 transistors of which three are defective. If three are chosen, in how many ways can one get two defective?
\end{puzzle}

\begin{puzzle}
    In how many ways can five counters labeled A, B, C, D and E at a store be staffed by two men and three women chosen from a group of four men and six women?
\end{puzzle}

\begin{puzzle}
    How many 4-letter word sequences consisting of two vowels and two consonants can be made from the letters of the word PHOENIX if no letter is repeated?
\end{puzzle}

Three marbles are chosen from an urn that contains 5 red, 4 white, and 3 blue marbles. How many samples
of the following type are possible?

\begin{puzzle}
    All three white.
\end{puzzle}

\begin{puzzle}
    Two blue and one white.
\end{puzzle}

\begin{puzzle}
    One of each color.
\end{puzzle}

\begin{puzzle}
    All three of the same color.
\end{puzzle}

\begin{puzzle}
    At least two red.
\end{puzzle}

\begin{puzzle}
    None red.
\end{puzzle}

Five coins are chosen from a bag that contains 4 dimes, 5 nickels, and 6 pennies. How many samples of five of the following type are possible?

\begin{puzzle}
    At least four nickels.
\end{puzzle}

\begin{puzzle}
    No pennies.
\end{puzzle}

\begin{puzzle}
    Five of a kind.
\end{puzzle}

\begin{puzzle}
    Four of a kind.
\end{puzzle}

\begin{puzzle}
    Two of one kind and two of another kind.
\end{puzzle}

\begin{puzzle}
    Three of one kind and two of another kind.
\end{puzzle}

Find the number of different ways to draw a 5-card hand from a deck to have the following combinations.

\begin{puzzle}
    Three face cards.
\end{puzzle}

\begin{puzzle}
    A heart flush (all hearts).
\end{puzzle}

\begin{puzzle}
    Two hearts and three diamonds.
\end{puzzle}

\begin{puzzle}
    Two cards of one suit, and three of another suit.
\end{puzzle}

\begin{puzzle}
    Two kings and three queens.
\end{puzzle}

\begin{puzzle}
    Two cards of one value and three of another value.
\end{puzzle}