\section{Introduction to Matrices}

\begin{center}
    \begin{tabular}{|l|cc|cc|}
        \hline
                  & January  &           & February &           \\
        \hline
                  & Hot Dogs & Corn Dogs & Hot Dogs & Corn Dogs \\
        \hline
        Place I   & 10       & 8         & 8        & 7         \\
        Place II  & 8        & 6         & 6        & 7         \\
        Place III & 6        & 4         & 6        & 5         \\
        \hline
    \end{tabular}
\end{center}

Represent these tables as \(3 \times 2\) matrices \(J\) and \(F\), and answer Puzzles \ref{puzzle_hotdog_sum}, \ref{puzzle_hotdog_difference}, \ref{puzzle_hotdog_revenue}, and \ref{puzzle_hotdog_increase}.


\begin{puzzle}\label{puzzle_hotdog_sum}
    Determine total sales for the two months, that is, find \(J + F\).
\end{puzzle}

\begin{puzzle}\label{puzzle_hotdog_difference}
    Find the difference in sales, \(J - F\).
\end{puzzle}

\begin{puzzle}\label{puzzle_hotdog_revenue}
    If hot dogs sell for \$3 and corn dogs for \$2, find the revenue from the sale of hot dogs and corn dogs. Hint: Let \(P\) be a \(2 \times 1\) matrix. Find \((J + F) P\).
\end{puzzle}

\begin{puzzle}\label{puzzle_hotdog_increase}
    If March sales will be up from February by 10\%, 15\%, and 20\% at Place I, Place II, and Place III, respectively, find the total expected number of hot dogs and corn dogs to be sold in March across all three places. Hint: Let \(R\) be a \(1 \times 3\) matrix with entries 1.10, 1.15, and 1.20. Your answer will be a 1 by 2 matrix.
\end{puzzle}

Determine the sums and products in the next 5 problems. Given the matrices \( A \), \( B \), \( C \), and \( D \) as follows:
\[
    A =
    \begin{bmatrix}
        3 & 6 & 1 \\
        0 & 1 & 3 \\
        2 & 4 & 1
    \end{bmatrix}
    \quad
    B =
    \begin{bmatrix}
        1 & -1 & 2 \\
        1 & 4  & 2 \\
        3 & 1  & 1
    \end{bmatrix}
    \quad
    C =
    \begin{bmatrix}
        1 \\
        2 \\
        3
    \end{bmatrix}
    \quad
    D =
    \begin{bmatrix}
        2 & 3 & 2
    \end{bmatrix}
\]

\begin{puzzle}
    $3A - 2B$
\end{puzzle}

\begin{puzzle}
    $AB + BA$
\end{puzzle}

\begin{puzzle}
    $A^2$
\end{puzzle}

\begin{puzzle}
    $2BC$
\end{puzzle}

\begin{puzzle}
    $2CD + 3AB$
\end{puzzle}

\begin{puzzle}
    $A^2B$
\end{puzzle}

\begin{puzzle}
    Let \( E = \begin{bmatrix} m & n \\ p & q \end{bmatrix} \) and \( F = \begin{bmatrix} a & b \\ c & d \end{bmatrix} \), find $EF$.
\end{puzzle}

\begin{puzzle}
    Let \( G = \begin{bmatrix}
        3 & 6 & 1 \\
        0 & 1 & 3 \\
        2 & 4 & 1 \\
    \end{bmatrix} \) and \( H = \begin{bmatrix}
        x \\
        y \\
        z \\
    \end{bmatrix} \), find GH.
\end{puzzle}

Express the following system as \( AX = B \), where \( A \), \( X \), and \( B \) are matrices.

\begin{puzzle}

    \[
        \begin{aligned}
            4x - 5y & = 6 \\
            5x - 6y & = 7
        \end{aligned}
    \]
\end{puzzle}

\begin{puzzle}
    \[
        \begin{aligned}
            x - 2y + 2z & = 3   \\
            x - 3y + 4z & = 7   \\
            x - 2y - 3z & = -12 \\
        \end{aligned}
    \]
\end{puzzle}

\begin{puzzle}
    \begin{align*}
        2x + 3z & = 17 \\
        3x - 2y & = 10 \\
        5y + 2z & = 11
    \end{align*}
\end{puzzle}

\begin{puzzle}
    \begin{align*}
        x + 2y + 3z + 2w & = 14 \\
        x - 2y - z       & = -5 \\
        y - 2z + 4w      & = 9  \\
        x + 3z + 3w      & = 15
    \end{align*}
\end{puzzle}