\section{Independent Events Puzzles}

The distribution of the number of fiction and non-fiction books checked out at a city's main library and at
a smaller branch on a given day is as follows.

\begin{center}
    \begin{tabular}{|l|c|c|c|}
        \hline
                        & MAIN (M) & BRANCH (B) & TOTAL \\ \hline
        FICTION (F)     & 300      & 100        & 400   \\
        \hline
        NON-FICTION (N) & 150      & 50         & 200   \\
        \hline
        TOTALS          & 450      & 150        & 600   \\
        \hline
    \end{tabular}
\end{center}
Use this table to determine the following probabilities:

\begin{puzzle}
    \(P(F)\)
\end{puzzle}

\begin{puzzle}
    \(P(M|F)\)
\end{puzzle}

\begin{puzzle}
    \(P(N|B)\)
\end{puzzle}

\begin{puzzle}
    Is the fact that a person checks out a fiction book independent of the main library?
\end{puzzle}

Toss two coins, let the events \( E \), \( F \), and \( G \) be as follows.
\[ E : \text{There is at least one head} \]
\[ F : \text{Both heads and tails are showing} \]
\[ G : \text{The first toss is a head} \]
\begin{puzzle}
    Find the following.
    \begin{enumerate}
        \item \( P(E) \)
        \item \( P(F) \)
        \item \( P(E \cap F) \)
        \item Are \( E \) and \( F \) independent?
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    Find the following.
    \begin{enumerate}
        \item \( P(F) \)
        \item \( P(G) \)
        \item \( P(F \cap G) \)
        \item Are \( F \) and \( G \) independent?
    \end{enumerate}
\end{puzzle}

Do the following problems involving independence.

\begin{puzzle}
    If \( P(E) = .6 \), \( P(F) = .2 \), and \( E \) and \( F \) are independent, find \( P(E \cap F) \).
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .6 \), \( P(F) = .2 \), and \( E \) and \( F \) are independent, find \( P(E \cup F) \).
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .9 \), \( P(F \mid E) = .36 \), and \( E \) and \( F \) are independent, find \( P(F) \).
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .6 \), \( P(E \cup F) = .08 \), and \( E \) and \( F \) are independent, find \( P(F) \).
\end{puzzle}

\begin{puzzle}
    In a survey of 100 people, 40 were casual drinkers, and 60 did not drink. Of the ones who drank, 6 had minor headaches. Of the non-drinkers, 9 had minor headaches. Are the events "drinkers" and "had headaches" independent?
\end{puzzle}

\begin{puzzle}
    It is known that 80\% of the people wear seat belts, and 5\% of the people quit smoking last year. If 4\% of the people who wear seat belts quit smoking, are the events, wearing a seat belt and quitting smoking, independent?
\end{puzzle}

\begin{puzzle}
    John's probability of passing statistics is 40\%, and Linda's probability of passing the same course is 70\%. If the two events are independent, find the following probabilities:
    \begin{enumerate}
        \item \( P(\text{both of them will pass statistics}) \)
        \item \( P(\text{at least one of them will pass statistics}) \)
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    Jane is flying home for the Christmas holidays. She has to change planes twice on the way home. There is an 80\% chance that she will make the first connection, and a 90\% chance that she will make the second connection. If the two events are independent, find the following probabilities:
    \begin{enumerate}
        \item \( P(\text{Jane will make both connections}) \)
        \item \( P(\text{Jane will make at least one connection}) \)
    \end{enumerate}
\end{puzzle}
