\section{Complex Perumation Puzzles}

Do the following problems using the techniques learned in this section.

\begin{puzzle}
    In how many different ways can five children hold hands to play "Ring Around the Rosy"?
    \end{puzzle}
    
    \begin{puzzle}
    In how many ways can three people be made to sit at a round table?
    \end{puzzle}
    
    \begin{puzzle}
    In how many different ways can six children ride a "Merry Go Around" with six horses?
    \end{puzzle}
    
    \begin{puzzle}
        In how many ways can three couples be seated at a round table, so that men and women sit alternately?
        \end{puzzle}
        
        \begin{puzzle}
        In how many ways can six trinkets be arranged on a chain?
        \end{puzzle}
        
        \begin{puzzle}
        In how many ways can five keys be put on a key ring?
        \end{puzzle}
        
        \begin{puzzle}
        Find the number of different permutations of the letters of the word MASSACHUSETTS.
        \end{puzzle}
        
        \begin{puzzle}
        Find the number of different permutations of the letters of the word MATHEMATICS.
        \end{puzzle}
        
        \begin{puzzle}
        Seven flags, three red, two white, and two blue, are to be flown on seven poles. How many different arrangements are possible?
        \end{puzzle}
        
        \begin{puzzle}
        How many different ways can three pennies, two nickels and five dimes be arranged in a row?
        \end{puzzle}

        \begin{puzzle}
            How many four-digit numbers can be made using two 2's and two 3's?
            \end{puzzle}
            
            \begin{puzzle}
            How many five-digit numbers can be made using two 6's and three 7's?
            \end{puzzle}
            
            \begin{puzzle}
            If a coin is tossed 5 times, how many different outcomes of 3 heads, and 2 tails are possible?
            \end{puzzle}
            
            \begin{puzzle}
            If a coin is tossed 10 times, how many different outcomes of 7 heads, and 3 tails are possible?
            \end{puzzle}
            
            \begin{puzzle}
            If a team plays ten games, how many different outcomes of 6 wins, and 4 losses are possible?
            \end{puzzle}
            
            \begin{puzzle}
            If a team plays ten games, how many different ways can the team have a winning season?
            \end{puzzle}
            