\section{Chapter Review Puzzles}

\begin{puzzle}
    The value of a new boat depreciates after it is purchased. The boat was bought for \$25,000 and depreciates at 8.2\% per year.
    \begin{enumerate}
        \item Write the function that models the value of the boat over time.
        \item Determine how many years until the boat's value decreases to \$20,000.
        \item What will be the value of the boat three years after it was purchased?
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    Tony invested \$40,000 in 2010, but the investment has been depreciating by 2.7\% per year.
    \begin{enumerate}
        \item Write the function representing the value of Tony's investment over time.
        \item Calculate the investment's value in 2020.
        \item Find out in which year the investment will be worth half of its original value.
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    Rosa invested \$25,000 in 2005, which has been appreciating by 6.4\% annually.
    \begin{enumerate}
        \item Write the formula for Rosa's investment value over time.
        \item Compute the value of the investment in 2025.
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    The population of a city has been increasing at 3.2\% annually since 2000. The population was 235,000 in 2015.
    \begin{enumerate}
        \item Find the population in the year 2000.
        \item Determine the year when the population will reach 250,000.
        \item Calculate the population in the year 2008.
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    An endangered species had a population of 5000, decreasing by 12\% annually.
    \begin{enumerate}
        \item What will the population be four years from now?
        \item In which year will the population decrease to only 2000 animals?
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    A patient receives a 300 mg dose of medication. After 5 hours, only 80 mg remains in the bloodstream.
    \begin{enumerate}
        \item Using an exponential decay model, determine the hourly decay rate.
        \item Calculate the time after the dose when only 125 mg remains in the bloodstream.
        \item Find out how much medication remains in the bloodstream after 8 hours.
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    If a relationship is described by \( y = 240t \) and \( y = 600 \) when \( t = 6 \) years, find the annual growth rate in percent.
\end{puzzle}

\begin{puzzle}
    Rewrite the given functions into the specified forms:
    \begin{enumerate}
        \item For \( y = 375000(1.125^t) \), convert it into the form \( y = ab^t \).
        \item For \( y = 5400e^{0.12t} \), express it as \( y = ae^{kt} \).
        \item For \( y = 230e^{-0.62t} \), represent it as \( y = ab^t \).
        \item For \( y = 3600(0.42^t) \), change it to \( y = ae^{kt} \).
    \end{enumerate}
\end{puzzle}

