\section{Simple Interest Puzzles}

Do the following simple interest problems:

\begin{puzzle}
    If an amount of \$2,000 is borrowed at a simple interest rate of 10\% for 3 years, how much is the interest?
\end{puzzle}

\begin{puzzle}
    You borrow \$4,500 for six months at a simple interest rate of 8\%. How much is the interest?
\end{puzzle}

\begin{puzzle}
    John borrows \$2400 for 3 years at 9\% simple interest. How much will he owe at the end of 3 years?
\end{puzzle}

\begin{puzzle}
    Jessi takes a loan of \$800 for 4 months at 12\% simple interest. How much does she owe at the end of the 4-month period?
\end{puzzle}

\begin{puzzle}
    If an amount of \$2,160, which includes a 10\% simple interest for 2 years, is paid back, how much was borrowed 2 years earlier?
\end{puzzle}

\begin{puzzle}
    Jamie just paid off a loan of \$2,544, which includes the principal and simple interest. If he took out the loan six months ago at 12\% simple interest, what was the amount borrowed?
\end{puzzle}

\begin{puzzle}
    Shanti charged \$800 on her charge card and did not make a payment for six months. If there is a monthly charge of 1.5\%, how much does she owe?
\end{puzzle}

\begin{puzzle}
    A credit card company charges 18\% interest on the unpaid balance. If you owed \$2000 three months ago and have been delinquent since, how much do you owe?
\end{puzzle}

\begin{puzzle}
    An amount of \$2000 is borrowed for 3 years. At the end of the three years, \$2660 is paid back. What was the simple interest rate?
\end{puzzle}

\begin{puzzle}
    Nancy borrowed \$1,800 and paid back \$1,920, four months later. What was the simple interest rate?
\end{puzzle}

\begin{puzzle}
    Jose agrees to pay \$2,000 in one year at an interest rate of 12\%. The bank subtracts the discount of 12\% of \$2,000, and gives the rest to Jose. Find the amount of the discount and the proceeds to Jose.
\end{puzzle}

\begin{puzzle}
    Tasha signs a note for a discounted loan agreeing to pay \$1200 in 8 months at an 18\% discount rate. Determine the amount of the discount and the proceeds to her.
\end{puzzle}

\begin{puzzle}
    An amount of \$8,000 is borrowed at a discount rate of 12\%, find the proceeds if the length of the loan is 7 months.
\end{puzzle}

\begin{puzzle}
    An amount of \$4,000 is borrowed at a discount rate of 10\%, find the proceeds if the length of the loan is 180 days.
\end{puzzle}

\begin{puzzle}
    Derek needs \$2400 new equipment for his shop. He can borrow this money at a discount rate of 14\% for a year. Find the amount of the loan he should ask for so that his proceeds are \$2400.
\end{puzzle}

\begin{puzzle}
    Mary owes June \$750, and wants to pay her off. She decides to borrow the amount from her bank at a discount rate of 16\%. If she borrows the money for 10 months, find the amount of the loan she should ask for so that her proceeds are \$750?
\end{puzzle}