\section{Mutually Exclusive Events and the Addition Rule Puzzles}

Determine whether the following pair of events are mutually exclusive.
\begin{puzzle}
    \[ A = \{\text{A person earns more than \$75,000 per year}\}\]
    \[B = \{\text{A person earns less than \$70,000} per year\}\]
\end{puzzle}

\begin{puzzle}
    A card is drawn from a deck.
    \[ C = \{\text{It is a King}\}\]
    \[D = \{\text{It is a heart}\}.\]
\end{puzzle}

\begin{puzzle}
    A die is rolled.
    \[ E = \{\text{An even number shows}\}\]
    \[F = \{\text{A number greater than 3 shows}\}\]
\end{puzzle}

\begin{puzzle}
    Two dice are rolled.
    \[ G = \{\text{The sum of dice is 8}\}\]
    \[H = \{\text{One die shows a 6}\}\]
\end{puzzle}

\begin{puzzle}
    Three coins are tossed.
    \[I = \{\text{Two heads come up}\}\]
    \[J = \{\text{At least one tail comes up}\}\]
\end{puzzle}

\begin{puzzle}
    Three coins are tossed.
    \[K = \{\text{First toss is a head}\}\]
    \[L = \{\text{both heads and tails appear}\}\]
\end{puzzle}

\begin{puzzle}
    A card is drawn from a deck, and the events \( C \) and \( D \) are as follows:
    \[ C = \{\text{It is a king}\}\]
    \[ D = \{\text{It is a heart}\}\]
    Find \( P(C \text{ or } D) \).
\end{puzzle}

\begin{puzzle}
    A die is rolled, and the events \( E \) and \( F \) are as follows:
    \[ E = \{\text{An even number shows}\}\]
    \[ F = \{\text{A number greater than 3 shows}\}\]
    Find \( P(E \text{ or } F) \).
\end{puzzle}

\begin{puzzle}
    Two dice are rolled, and the events \( G \) and \( H \) are as follows:
    \[ G = \{\text{The sum of dice is 8}\}\]
    \[ H = \{\text{Exactly one die shows a 6}\}\]
    Find \( P(G \text{ or } H) \).
\end{puzzle}

\begin{puzzle}
    Three coins are tossed, and the events \( I \) and \( J \) are as follows:
    \[ I = \{\text{Two heads come up}\}\]
    \[ J = \{\text{At least one tail comes up}\}\]
    Find \( P(I \text{ or } J) \).
\end{puzzle}


\begin{puzzle}
    At De Anza college, 20\% of the students take Finite Mathematics, 30\% take Statistics and 10\% take both. What percentage of the students take Finite Mathematics or Statistics?
\end{puzzle}

\begin{puzzle}
    This quarter, there is a 50\% chance that Jason will pass Accounting, a 60\% chance that he will
    pass English, and 80\% chance that he will pass at least one of these two courses. What is the
    probability that he will pass both Accounting and English?
\end{puzzle}

The following table shows the distribution of Democratic and Republican U.S. Senators by gender.
\begin{center}
    \begin{tabular}{|l|c|c|c|c|}
        \hline
                        & Men (M) & Women (F) & Nonbinary (N) & Total \\ \hline
        Democrat (D)    & 33      & 15        & 0             & 48    \\
        Republican (R)  & 40      & 9         & 0             & 49    \\
        Independent (I) & 2       & 1         & 0             & 3     \\
        Total           & 75      & 25        & 0             & 100   \\ \hline
    \end{tabular}
\end{center}

\begin{puzzle}
    Find the probability of event \( M \) and event \( D \) occurring together:
    \[ P(M \text{ and } D) \]
\end{puzzle}

\begin{puzzle}
    Find the probability of event \( F \) and event \( R \) occurring together:
    \[ P(F \text{ and } R) \]
\end{puzzle}

\begin{puzzle}
    Find the probability of event \( M \) or event \( D \) occurring:
    \[ P(M \text{ or } D) \]
\end{puzzle}

\begin{puzzle}
    Find the probability of event \( F \) or the complement of event \( R \) occurring:
    \[ P(F \text{ or } R^C) \]
\end{puzzle}

\begin{puzzle}
    Find the probability of the complement of event \( M \) or event \( R \) occurring:
    \[ P(M^C \text{ or } R) \]
\end{puzzle}

\begin{puzzle}
    Find the probability of event \( M \) or event \( F \) occurring:
    \[ P(M \text{ or } F) \]
\end{puzzle}

Again, use the addition rule to determine the following probabilities.

\begin{puzzle}
    If \( P(E) = .5 \) and \( P(F) = .4 \) and \( E \) and \( F \) are mutually exclusive, find \( P(E \text{ and } F) \).
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .4 \) and \( P(F) = .2 \) and \( E \) and \( F \) are mutually exclusive, find \( P(E \text{ or } F) \).
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .3 \) and \( P(E \text{ or } F) = .6 \) and \( P(E \text{ and } F) = .2 \), find \( P(F) \).
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .4 \), \( P(F) = .5 \) and \( P(E \text{ or } F) = .7 \), find \( P(E \text{ and } F) \).
\end{puzzle}
