\section{Compound Interest Puzzles}

Do the following compound interest problems involving a lump-sum amount:

\begin{puzzle}
    If \$8,000 is invested at 9.2\% compounded monthly, what will the final amount be in 4 years?
\end{puzzle}

\begin{puzzle}
    How much should be invested at 10.3\% for it to amount to \$10,000 in 6 years?
\end{puzzle}

\begin{puzzle}
    Lydia’s aunt Rose left her \$5,000. Lydia spent \$1,000 on her wardrobe and deposited the rest in an account that pays 6.9\% compounded daily. How much money will she have in 5 years?
\end{puzzle}

\begin{puzzle}
    Thuy needs \$1,850 in eight months for her college tuition. How much money should she deposit lump sum in an account paying 8.2\% compounded monthly to achieve that goal?
\end{puzzle}

\begin{puzzle}
    Bank A pays 5\% compounded daily, while Bank B pays 5.12\% compounded monthly. Which bank pays more? Explain.
\end{puzzle}

\begin{puzzle}
    EZ Photo Company needs five copying machines in 2 1/2 years for a total cost of \$15,000. How much money should be deposited now to pay for these machines if the interest rate is 8\% compounded semiannually?
\end{puzzle}

\begin{puzzle}
    Jon’s grandfather was planning to give him \$12,000 in 10 years. Jon has convinced his grandfather to pay him \$6,000 now, instead. If Jon invests this \$6,000 at 7.5\% compounded continuously, how much money will he have in 10 years?
\end{puzzle}

\begin{puzzle}
    What will be the price of a \$20,000 car in 5 years if the inflation rate is 6\%?
\end{puzzle}

\begin{puzzle}
    At an interest rate of 8\% compounded continuously, how many years will it take to double your money? Hint: You may do this on your calculator by trial and error.
\end{puzzle}

\begin{puzzle}
    If an investment earns 10\% compounded continuously, in how many years will it triple? Hint: You may do this on your calculator by trial and error.
\end{puzzle}

\begin{puzzle}
    The City Library has ordered a new computer system costing \$158,000. The system will be delivered in 6 months, and the full amount will be due 30 days after delivery. How much should be deposited today into an account paying 7.5\% compounded monthly to have \$158,000 in 7 months?
\end{puzzle}

\begin{puzzle}
    Mr. and Mrs. Tran are expecting a baby girl in a few days. They want to put away money for her college education now. How much money should they deposit in an account paying 10.2\% so they will have \$100,000 in 18 years to pay for their daughter’s educational expenses?
\end{puzzle}

\begin{puzzle}
    Find the effective interest rate for an account paying 7.2\% compounded quarterly.
\end{puzzle}

\begin{puzzle}
    If a bank pays 5.75\% compounded monthly, what is the effective interest rate?
\end{puzzle}

\begin{puzzle}
    Population of California in the year 1995 was 32 million. If the population grows at a rate of 2\%, what will the population be in 2025?
\end{puzzle}

\begin{puzzle}
    According to the Law of 70, if an amount grows at an annual rate of 1\%, then it doubles every seventy years. Suppose a bank pays 5\% interest, how long will it take for you to double your money? How about at 15\%?
\end{puzzle}
