\section{Perumations Puzzles}

Do the following problems using permutations.

\begin{puzzle}
    How many three-letter words can be made using the letters \{a, b, c, d, e, f\} if no repetitions are allowed?
\end{puzzle}

\begin{puzzle}
    A grocery store has five checkout counters, and seven clerks. How many different ways can the clerks be assigned to the counters?
\end{puzzle}

\begin{puzzle}
    A group of fifteen people who are members of an investment club wish to choose a president, and a secretary. How many different ways can this be done?
\end{puzzle}

\begin{puzzle}
    Compute the following.
    \begin{enumerate}
        \item \( P(9,2) \)
        \item \( P(6,4) \)
        \item \( P(8,3) \)
        \item \( P(7,4) \)
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    In how many ways can the letters of the word CUPERTINO be arranged if each letter is used only once in each arrangement?
\end{puzzle}

\begin{puzzle}
    How many permutations of the letters of the word PROBLEM end in a vowel?
\end{puzzle}

\begin{puzzle}
    How many permutations of the letters of the word SECURITY end in a consonant?
\end{puzzle}

\begin{puzzle}
    How many permutations of the letters PRODUCT have consonants in the second and third positions?
\end{puzzle}

\begin{puzzle}
    How many three-digit numbers are there?
\end{puzzle}

\begin{puzzle}
    How many three-digit odd numbers are there?
\end{puzzle}

\begin{puzzle}
    In how many different ways can five people be seated in a row if two of them insist on sitting next to each other?
\end{puzzle}

\begin{puzzle}
    In how many different ways can five people be seated in a row if two of them insist on not sitting next to each other?
\end{puzzle}

\begin{puzzle}
    In how many ways can 3 English, 3 history, and 2 math books be set on a shelf, if the English books are set on the left, history books in the middle, and math books on the right?
\end{puzzle}

\begin{puzzle}
    In how many ways can 3 English, 3 history, and 2 math books be set on a shelf, if they are grouped by subject?
\end{puzzle}

\begin{puzzle}
    You have 5 math books and 6 history books to put on a shelf with five slots. In how many ways can you put the books on the shelf if the first two slots are to be filled with math books and the next three with history books?
\end{puzzle}

\begin{puzzle}
    You have 5 math books and 6 history books to put on a shelf with five slots. How many ways can you put the books on the shelf if the first two slots are to be filled with books of the same subject and the next three slots are to be filled with books of the other subject?
\end{puzzle}
