\section{Annuities Puzzles}

Each of the following problems involve an annuity - a sequence of payments.

\begin{puzzle}
    Find the future value of an annuity of \$200 per month for 5 years at 6\% compounded monthly.
    \end{puzzle}
    
    \begin{puzzle}
    How much money should be deposited at the end of each month in an account paying 7.5\% for it to amount to \$10,000 in 5 years?
    \end{puzzle}
    
    \begin{puzzle}
    At the end of each month Rita deposits \$300 in an account that pays 5\%. What will the final amount be in 4 years?
    \end{puzzle}
    
    \begin{puzzle}
    Mr. Chang wants to retire in 10 years and can save \$650 every three months. If the interest rate is 7.8\%, how much will he have at the end of 5 years?
    \end{puzzle}
    
    \begin{puzzle}
    A firm needs to replace most of its machinery in five years at a cost of \$500,000. The company wishes to create a sinking fund to have this money available in five years. How much should the quarterly deposits be if the fund earns 8\%?
    \end{puzzle}
    
    \begin{puzzle}
    Mrs. Brown needs \$5,000 in three years. If the interest rate is 9\%, how much should she save at the end of each month to have that amount in three years?
    \end{puzzle}
    
    \begin{puzzle}
        A company has a \$120,000 note due in 4 years. How much should be deposited at the end of each quarter in a sinking fund to payoff the note in four years if the interest rate is 8\%?
        \end{puzzle}
        
        \begin{puzzle}
        You are now 20 years of age and decide to save \$100 at the end of each month until you are 65. If the interest rate is 9.2\%, how much money will you have when you are 65?
        \end{puzzle}
        
        \begin{puzzle}
        Is it better to receive \$400 at the beginning of each month for six years, or a lump sum of \$25,000 today if the interest rate is 7\%? Explain.
        \end{puzzle}
        
        \begin{puzzle}
        In order to save money for a new computer Jill decided to save \$125 at the beginning of each month for the next 8 months. If the interest rate is 7\%, how much money will she have at the end of 8 months?
        \end{puzzle}
        
        \begin{puzzle}
        Mrs. Gill puts \$2200 at the end of each year in her IRA account that earns 9\% per year. How much total money will she have in this account after 20 years?
        \end{puzzle}
        
        \begin{puzzle}
        If the inflation rate stays at 6\% per year for the next five years, how much will the price be of a \$15,000 car in five years? How much must you save at the end of each month at an interest rate of 7.3\% to buy that car in 5 years?
        \end{puzzle}
        