\section{Minimization Puzzles}

For each of the following minimization problems, choose your variables, write the objective function and the
constraints, graph the constraints, shade the feasibility region, label all critical points, and determine the
solution that optimizes the objective function

\begin{puzzle}
    A diet is to contain at least 2400 units of vitamins, 1800 units of minerals, and 1200 calories. Two
    foods, Food A and Food B are to be purchased. Each unit of Food A provides 50 units of vitamins,
    30 units of minerals, and 10 calories. Each unit of Food B provides 20 units of vitamins, 20 units
    of minerals, and 40 calories. If Food A costs \$2 per unit and Food B cost \$1 per unit, how many
    units of each food should be purchased to keep costs at a minimum?
\end{puzzle}

\begin{puzzle}
    A computer store sells two types of computers, desktops and laptops. The supplier demands that
    at least 150 computers be sold a month. In order to keep profits up, the number of desktops sold
    must be at least twice of laptops. The store pays its sales staff a \$75 commission for each desk
    top, and a \$50 commission for each lap top. How many of each type of computers must be sold to
    minimize commission to its sales people? What is the minimum commission?
\end{puzzle}

\begin{puzzle}
    An oil company has two refineries. Each day, Refinery A produces 200 barrels of high-grade oil,
    300 barrels of medium-grade oil, and 200 barrels of low-grade oil and costs \$12,000 to operate.
    Each day, Refinery B produces 100 barrels of high-grade oil, 100 barrels of medium-grade oil, and
    200 barrels of low-grade oil and costs \$10,000 to operate. The company must produce at least 800
    barrels of high-grade oil, 900 barrels of medium-grade oil, and 1,000 barrels of low-grade oil. How
    many days should each refinery be operated to meet the goals at a minimum cost?

\end{puzzle}

\begin{puzzle}
    A print shop at a community college in Cupertino, California, employs two different contractors
    to maintain its copying machines. The print shop needs to have 12 IBM, 18 Xerox, and 20 Canon
    copying machines serviced. Contractor A can repair 2 IBM, 1 Xerox, and 2 Canon machines at a
    cost of \$800 per month, while Contractor B can repair 1 IBM, 3 Xerox, and 2 Canon machines at
    a cost of \$1000 per month. How many months should each of the two contractors be employed to
    minimize the cost?
\end{puzzle}
