\section{Conditional Probability Puzzles}

Do the following puzzles using the conditional probability formula: \[ P(A | B) = \frac{P(A \cap B)}{P(B)}. \]

\begin{puzzle}
    A card is drawn from a deck. Find the conditional probability of $P(\text{a queen} | \text{a face card})$.
\end{puzzle}

\begin{puzzle}
    A card is drawn from a deck. Find the conditional probability of $P(\text{a queen} | \text{a club})$.
\end{puzzle}

\begin{puzzle}
    A die is rolled. Find the conditional probability that it shows a three if it is known that an odd number has shown.
\end{puzzle}

\begin{puzzle}
    If $P(A) = .3$ and $P(B) = .4$, and $P(A \cap B) = .12$, find the following:
    \begin{enumerate}
        \item $P(A | B)$
        \item $P(B | A)$
    \end{enumerate}
\end{puzzle}



The following table shows the distribution of Democratic and Republican U.S. Senators by gender.
\begin{center}
    \begin{tabular}{|l|c|c|c|c|}
        \hline
                        & Men (M) & Women (F) & Nonbinary (N) & Total \\ \hline
        Democrat (D)    & 33      & 15        & 0             & 48    \\
        Republican (R)  & 40      & 9         & 0             & 49    \\
        Independent (I) & 2       & 1         & 0             & 3     \\
        Total           & 75      & 25        & 0             & 100   \\ \hline
    \end{tabular}
\end{center}

Use this table to determine the following probabilities:

\begin{puzzle}
    \( P(M | D) \)
\end{puzzle}

\begin{puzzle}
    \( P(D | M) \)
\end{puzzle}

\begin{puzzle}
    \( P(F | R) \)
\end{puzzle}

\begin{puzzle}
    \( P(R | F) \)
\end{puzzle}

Do the following conditional probability problems.

\begin{puzzle}
    At De Anza College, 20\% of the students take Finite Math, 30\% take History, and 5\% take both Finite Math and History. If a student is chosen at random, find the following conditional probabilities.
    \begin{enumerate}
        \item They are taking Finite Math given that they are taking History
        \item They are taking History assuming that they are taking Finite Math.
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    At a college, 60\% of the students pass Accounting, 70\% pass English, and 30\% pass both of these courses. If a student is selected at random, find the following conditional probabilities.
    \begin{enumerate}
        \item They pass Accounting given that they passed English
        \item They pass English assuming that they passed Accounting.
    \end{enumerate}
\end{puzzle}


\begin{puzzle}
    If \( P(F') = .4 \) and \( P(E' | F) = .3 \), find \( P(E' \text{ and } F) \).
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .3 \), and \( P(F) = .3 \), and \( E \) and \( F \) are mutually exclusive, find \( P(E | F') \).
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .6 \) and \( P(E \text{ and } F) = .24 \), find \( P(F | E) \).
\end{puzzle}

\begin{puzzle}
    If \( P(E \text{ and } F) = .04 \) and \( P(E | F') = .1 \), find \( P(F) \).
\end{puzzle}



Consider three successive coin tosses. Find the following probabilities.

\begin{puzzle}
    \( P(\text{two heads } | \text{ first toss is a head})\)
\end{puzzle}

\begin{puzzle}
    \( P(\text{all tails } | \text{ at least one toss is tails})\)
\end{puzzle}

\begin{puzzle}
    \( P(\text{both heads and tails } | \text{ first toss is heads})\)
\end{puzzle}

\begin{puzzle}
    \( P(\text{all heads } | \text{ there are both heads and tails})\)
\end{puzzle}