\section{Expected Value Puzzles}

\begin{puzzle}
    You are about to make an investment which gives you a 30\% chance of making \$60,000 and 70\% chance of losing \$30,000. Should you invest? Explain.
\end{puzzle}

\begin{puzzle}
    At Columbia college, 40\% of the staff and 30\% of the faculty are taller than 5'8''. If the college employees are 54\% staff and 46\% faculty, what percent of them are taller than 5'8''?
\end{puzzle}

\begin{puzzle}
    A game involves rolling a four sided die (1d4). If a one, two, or three shows, the player receives the face value of the die in dollars, but if a four shows, the player is obligated to pay \$4. What is the expected value of the game?
\end{puzzle}

\begin{puzzle}
    A game involves rolling a single die. One receives the face value of the die in dollars. How much should one be willing to pay to roll the die to make the game fair?
\end{puzzle}

\begin{puzzle}
    In a European country, 20\% of the families have three children, 40\% have two children, 30\% have one child, and 10\% have no children. On average, how many children are there to a family?
\end{puzzle}

\begin{puzzle}
    A game involves drawing a single card from a standard deck. One receives 60 cents for an ace, 30 cents for a king, and 5 cents for a red card that is neither an ace nor a king. If the cost of each draw is 10 cents, should one play? Explain.
\end{puzzle}

\begin{puzzle}
    FFA plans to raise money by raffling a saddle worth \$500. A total of 3000 tickets are sold at \$1 each. Find the expected value of the winnings for a person who buys a ticket in the raffle.
\end{puzzle}

\begin{puzzle}
    During her four years at college, Niki received A's in 30\% of her units, B's in 60\% of her units, and C's in the remaining 10\%. If A = 4, B = 3, and C = 2, find her grade point average. (Notice, GPA is calculated using units not using classes. Getting an C in a 5 unit class and an A in a 2 unit class results in a GPA of 2.57, while getting an A in the 5 unit class and a C in the 2 unit class results in a GPA of 3.43.)
\end{puzzle}

\begin{puzzle}
    Attendance at a Stanford football game depends upon which team Stanford is playing against. If the game is against U. C. Berkeley, the attendance will be 70,000; if it is against another California team, it will be 40,000; and if it is against an out of state team, it will be 30,000. If the probability of playing against U. C. Berkeley is 10\%, against a California team 50\%, and against an out of state team 40\%, how many fans are expected to attend a game?
\end{puzzle}

\begin{puzzle}
    A Texas oil drilling company has determined that it costs \$25,000 to sink a test well. If oil is hit, the revenue for the company will be \$500,000. If natural gas is found, the revenue will be \$150,000. If the probability of hitting oil is 3\% and of hitting gas is 6\%, find the expected value of sinking a test well.
\end{puzzle}

\begin{puzzle}
    A \$1 lottery ticket offers a grand prize of \$10,000; 10 runner-up prizes each paying \$100; and 1,000 fourth-place prizes each paying \$1. Find the expected value of entering this contest if 1 million tickets are sold.
\end{puzzle}

\begin{puzzle}
    Assume that for the next heavyweight fight the odds of Mike Tyson winning are 15 to 2. If a person bets \$10 that Mike Tyson will lose. If Mike Tyson loses, how much should the person collect?
\end{puzzle}


