\section{Game Theory Puzzles}

\begin{puzzle}
    Determine whether the games are strictly determined. If the games are strictly determined, find the optimal strategies for each player and the value of the game.
    \begin{enumerate}
        \item \[
                  \begin{bmatrix} 1 & 2 \\ -2 & 3 \end{bmatrix}
              \]
        \item \[
                  \begin{bmatrix} 6 & 3 \\ 2 & 1 \end{bmatrix}
              \]
        \item \[
                  \begin{bmatrix} -1 & -3 & 2 \\ 0 & 3 & -1 \\ 1 & -2 & 4 \\ 2 & 0 & -4 \end{bmatrix}
              \]
        \item \[
                  \begin{bmatrix} 3 & 4 & 2 \\ 0 & -2 & -3 \\ 0 & 2 & 1 \end{bmatrix}
              \]
        \item \[
                  \begin{bmatrix} -1 & -1 \\ -1 & 1 \\ 3 & 2 \end{bmatrix}
              \]
        \item \[
                  \begin{bmatrix} 5 & -3 & 2 \\ 3 & -1 & 4 \end{bmatrix}
              \]

    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    Two players play a game which involves holding out one or two fingers simultaneously. If the sum of the fingers is more than 2, Player II pays Player I the sum of the fingers; otherwise, Player I pays Player II the sum of the fingers.
    \begin{enumerate}
        \item Write a payoff matrix for Player I.
        \item Find the optimal strategies for each player and the value of the game.
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    A mayor of a large city is thinking of running for re-election, but does not know who his opponent is going to be. It is now time for him to take a stand for or against abortion. If he comes out against abortion rights and his opponent is for abortion, he will increase his chances of winning by 10\%. But if he is against abortion and so is his opponent, he gains only 5\%. On the other hand, if he is for abortion and his opponent against, he decreases his chance by 8\%, and if he is for abortion and so is his opponent, he decreases his chance by 12\%.
    \begin{enumerate}
        \item Write a payoff matrix for the mayor.
        \item Find the optimal strategies for the mayor and his opponent.
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    A man accused of a crime is not sure whether anybody saw him do it. He needs to make a choice of pleading innocent or pleading guilty to a lesser charge. If he pleads innocent and nobody comes forth, he goes free. However, if a witness comes forth, the man will be sentenced to 10 years in prison. On the other hand, if he pleads guilty to a lesser charge and nobody comes forth, he gets a sentence of one year and if a witness comes forth, he gets a sentence of 3 years.
    \begin{enumerate}
        \item Write a payoff matrix for the accused.
        \item If you were his attorney, what strategy would you advise?
    \end{enumerate}
\end{puzzle}