\section{Tree Diagram Puzzles}

Do the following problems using a tree diagram or the multiplcation axiom.

\begin{puzzle}
    A man has 3 shirts, and 2 pairs of pants. Use a tree diagram to determine the number of possible outfits.
\end{puzzle}

\begin{puzzle}
    In a city election, there are 2 candidates for mayor, and 3 for supervisor. Use a tree diagram to find the number of ways to fill the two offices.
\end{puzzle}

\begin{puzzle}
    There are 4 roads from Town A to Town B, 2 roads from Town B to Town C. Use a tree diagram to find the number of ways one can travel from Town A to Town C.
\end{puzzle}

\begin{puzzle}
    Brown Home Construction offers a selection of 3 floor plans, 2 roof types, and 2 exterior wall types. Use a tree diagram to determine the number of possible homes available.
\end{puzzle}

\begin{puzzle}
    For lunch, a small restaurant offers 2 types of soups, three kinds of sandwiches, and two types of soft drinks. Use a tree diagram to determine the number of possible meals consisting of a soup, sandwich, and a soft drink.
\end{puzzle}

\begin{puzzle}
    A California license plate consists of a number from 1 to 5, then three letters followed by three digits. How many such plates are possible?
\end{puzzle}

\begin{puzzle}
    A license plate consists of three letters followed by three digits. How many license plates are possible if no letter may be repeated?
\end{puzzle}

\begin{puzzle}
    How many different 4-letter radio station call letters can be made if the first letter must be K or W and none of the letters may be repeated?
\end{puzzle}

\begin{puzzle}
    How many seven-digit telephone numbers are possible if the first two digits cannot be ones or zeros?
\end{puzzle}

\begin{puzzle}
    How many 3-letter word sequences can be formed using the letters \{a, b, c, d\} if no letter is to be repeated?
\end{puzzle}

\begin{puzzle}
    A family has two children, use a tree diagram to determine all four possibilities.
\end{puzzle}

\begin{puzzle}
    A coin is tossed three times and the sequence of heads and tails is recorded. Use a tree diagram to determine the different possibilities.
\end{puzzle}

\begin{puzzle}
    In how many ways can a 4-question true-false test be answered?
\end{puzzle}

\begin{puzzle}
    In how many ways can three people be made to stand in a straight line?
\end{puzzle}

\begin{puzzle}
    A combination lock is opened by first turning to the left, then to the right, and then to the left again. If there are 30 digits on the dial, how many possible combinations are there?
\end{puzzle}

\begin{puzzle}
    How many different answers are possible for a multiple-choice test with 10 questions and five possible answers for each question?
\end{puzzle}