\section{Combination Puzzles}

Do the following problems using combinations.

\begin{puzzle}
    How many different 3-people committees can be chosen from ten people?
\end{puzzle}

\begin{puzzle}
    How many different 5-player teams can be chosen from eight players?
\end{puzzle}

\begin{puzzle}
    In how many ways can a person choose to vote for three out of five candidates on a ballot for a school board election?
\end{puzzle}

\begin{puzzle}
    Compute the following:
    \begin{enumerate}
        \item \( 9C2 \)
        \item \( 6C4 \)
        \item \( 8C3 \)
        \item \( 7C4 \)
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    How many 5-card hands can be chosen from a deck of cards?
\end{puzzle}

\begin{puzzle}
    How many 13-card bridge hands can be chosen from a deck of cards?
\end{puzzle}

\begin{puzzle}
    There are twelve people at a party. If they all shake hands, how many different handshakes are there?
\end{puzzle}

\begin{puzzle}
    In how many ways can a student choose to do four questions out of five on a test?
\end{puzzle}

\begin{puzzle}
    Five points lie on a circle. How many chords can be drawn through them?
\end{puzzle}

\begin{puzzle}
    How many diagonals does a hexagon have?
\end{puzzle}

\begin{puzzle}
    There are five teams in a league. How many games are played if every team plays each other twice?
\end{puzzle}

\begin{puzzle}
    A team plays 15 games a season. In how many ways can it have 8 wins and 7 losses?
\end{puzzle}

\begin{puzzle}
    In how many different ways can a 4-child family have 2 boys and 2 girls?
\end{puzzle}

\begin{puzzle}
    A coin is tossed five times. In how many ways can it fall three heads and two tails?
\end{puzzle}

\begin{puzzle}\label{puzzle_taxicab_combinations}
    The shopping area of a town is a square that is six blocks by six blocks. How many different routes can a taxi driver take to go from one corner of the shopping area to the opposite corner?
\end{puzzle}

\begin{puzzle}
    If the shopping area in Puzzle \ref{puzzle_taxicab_combinations} has a rectangular form of 5 blocks by 3 blocks, then how many different routes can a taxi driver take to drive from one end of the shopping area to the opposite corner?
\end{puzzle}
