\section{Binomial Probability Puzzles}

\begin{puzzle}
    A coin is tossed ten times. Find the probability of getting six heads and four tails.
\end{puzzle}

\begin{puzzle}
    A coin is tossed three times. Find the probability of having one head and two tails.
\end{puzzle}

\begin{puzzle}
    What is the probability of getting three ones if a die is rolled five times?
\end{puzzle}

\begin{puzzle}
    A baseball player has a .250 batting average. What is the probability that he will have three hits in five times at bat?
\end{puzzle}

\begin{puzzle}
    A basketball player has an 80\% chance of sinking a basket on a free throw. What is the probability that he will sink at least three baskets in five free throws?
\end{puzzle}

\begin{puzzle}
    With a new flu vaccination, 85\% of the people in the high risk group can go through the entire winter without contracting the flu. In a group of six people who were vaccinated with this drug, what is the probability that at least four will not get the flu?
\end{puzzle}

\begin{puzzle}
    A transistor manufacturer has known that 5\% of the transistors produced are defective. What is the probability that a batch of twenty five will have two defective?
\end{puzzle}

\begin{puzzle}
    It has been determined that only 80\% of the people wear seat belts. If a police officer stops a car with four people, what is the probability that at least one person will not be wearing a seat belt?
\end{puzzle}

\begin{puzzle}
    What is the probability that  five tossed coins will have at least three heads?
\end{puzzle}

\begin{puzzle}
    What is the probability that a toss of four coins will yield at most two heads?
\end{puzzle}
