\section{Application of Matrices in Cryptography}

In the following problems, the letters A to Z correspond to the numbers 1 to 26, as shown below, and a space is represented by the number 27.

\begin{center}
    \begin{tabular}{cccccccccccccc}
        A & B & C & D & E & F & G & H & I & J  & K  & L  & M  \\
        1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 \\
    \end{tabular}

    \begin{tabular}{cccccccccccccc}
        N  & O  & P  & Q  & R  & S  & T  & U  & V  & W  & X  & Y  & Z  \\
        14 & 15 & 16 & 17 & 18 & 19 & 20 & 21 & 22 & 23 & 24 & 25 & 26 \\
    \end{tabular}
\end{center}

In the next two problems, use the matrix \( A \), given below, to encode the given messages.

\[ A = \begin{bmatrix}
        3 & 2 \\
        1 & 1 \\
    \end{bmatrix} \]

In the two problems following, decode the messages that were encoded using matrix \( A \).
Make sure to consider the spaces between words, but ignore all punctuation. Add a final space if necessary.


\begin{puzzle}
    Encode the message: WATCH OUT!
\end{puzzle}

\begin{puzzle}
    Encode the message: HELP IS ON THE WAY.
\end{puzzle}

\begin{puzzle}
    Decode the following message:
    64 23 102 41 82 32 97 35 71 28 69 32
\end{puzzle}

\begin{puzzle}
    Decode the following message:
    105 40 117 48 39 19 69 32 72 27 37 15 114 47
\end{puzzle}

In the next two problems, use the matrix $B$, given below, to encode the given messages.
\[
    B = \begin{bmatrix}
        1 & 0 & 0  \\
        2 & 1 & 2  \\
        1 & 0 & -1 \\
    \end{bmatrix}
\]

Make sure to consider the spaces between words, but ignore all punctuation. Add a final space(s) if
necessary.

\begin{puzzle}
    Encode the message using matrix $B$:
    LUCK IS ON YOUR SIDE.
\end{puzzle}


\begin{puzzle}
    Encode the message using matrix $B$:
    MAY THE FORCE BE WITH YOU.
\end{puzzle}

In the two problems following, decode the messages that were encoded using matrix $B$.

\begin{puzzle}
    Decode the following message that was encoded using matrix B:
    \[
        8,~ 23,~ 7,~ 4,~ 47,~ -2,~ 15,~ 102,~ -12,~ 20,~ 58,~ 15,~ 27,~ 80,~ 18,~ 12,~ 74,~ -7
    \]
\end{puzzle}

\begin{puzzle}
    Decode the following message that was encoded using matrix B:
    \[
        12,~ 69,~ -3,~ 11,~ 53,~ 9,~ 46,~ -10,~ 18,~ 95,~ -9,~ 25,~ 107,~ 4,~ 27,~ 76,~ 22,~ 1,~ 72,~ -26
    \]
\end{puzzle}




