\section{Misc Application Puzzles}

For the following four problems, assume a \$200,000 house loan is amortized over 30 years at an interest rate
of 10.4\%.

\begin{puzzle}
    Find the monthly payment.
\end{puzzle}

\begin{puzzle}
    Find the balance owed after 20 years.
\end{puzzle}

\begin{puzzle}
    Find the balance of the loan after 100 payments.
\end{puzzle}

\begin{puzzle}
    Find the monthly payment if the original loan were amortized over 15 years.
\end{puzzle}

\begin{puzzle}
    Mr. Patel wants to pay off his car loan. The monthly payment for his car is \$365, and he has 16 payments left. If the loan was financed at 9\%, how much does he owe?
\end{puzzle}

\begin{puzzle}
    An amount of \$2000 is borrowed for a year at a rate of 18\%. Make an amortization schedule showing the monthly payment, the monthly interest on the outstanding balance, the portion of the payment going toward reducing the debt, and the balance.
\end{puzzle}

\begin{puzzle}
    Fourteen months after Dan bought his new car he lost his job. His car was repossessed by his lender after he had made only 14 monthly payments of \$376 each. If the loan was financed over a 4-year period at an interest rate of 10.3\%, how much did the car cost the lender? In other words, how much did Dan still owe on the car?
\end{puzzle}

\begin{puzzle}
    You have a choice of either receiving \$5,000 at the end of each year for the next 5 years or receiving \$3000 per year for the next 10 years. If the current interest rate is 9\%, which is better?
\end{puzzle}

\begin{puzzle}
    Mr. Smith is planning to retire in 25 years and would like to have \$250,000 then. What monthly payment made at the end of each month to an account that pays 6.5\% will achieve his objective?
\end{puzzle}

\begin{puzzle}
    Assume Mr. Smith has reached retirement and has \$250,000 in an account which is earning 6.5\%. He would now like to make equal monthly withdrawals for the next 15 years to completely deplete this account. Find the withdrawal payment.
\end{puzzle}

\begin{puzzle}
    A ten-year \$1,000 bond pays \$35 every six months. If the current interest rate is 8.2\%, find the fair market value of the bond.
    \begin{enumerate}
        \item Find the present value of \$1000.
        \item Find the present value of the \$35 payments.
        \item The fair market value of the bond \(=\) a \(+\) b.
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    Find the fair market value of the ten-year \$1,000 bond that pays \$35 every six months, if the current interest rate has dropped to 6\%.
    \begin{enumerate}
        \item Find the present value of \$1000.
        \item Find the present value of the \$35 payments.
        \item The fair market value of the bond \(=\) a \(+\) b.
    \end{enumerate}
\end{puzzle}
