\section{Growth and Decay Puzzles}

\begin{puzzle}
    Identify each as an exponential, linear, or power function:
    \begin{enumerate}
        \item \( y = 640 (1.25^x) \)
        \item \( y = 640 (x^{1.25}) \)
        \item \( y = 640 (1.25x) \)
        \item \( y = 1.05x - 2.5 \)
        \item \( y = 90 - \frac{4}{5}x \)
        \item \( y = 42(0.92^x) \)
        \item \( y = 37(x^{0.25}) \)
        \item \( y = 4(\frac{1}{3}^x) \)
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    Indicate if the function represents exponential growth or exponential decay:
    \begin{enumerate}
        \item \( y = 127e^{-0.35t} \)
        \item \( y = 70 (0.8^t) \)
        \item \( y = 453(1.2^t) \)
        \item \( y = 16e^{0.2t} \)
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    In each of the following, \( y \) is an exponential function of \( t \) stated in the form \( y = ae^{kt} \) where \( t \) represents time measured in years. For each:
    \begin{itemize}
        \item Re-express each function in the form \( y = ab^t \) (state the value of \( b \) accurate to 4 decimal places)
        \item State the annual growth rate or annual decay rate as a percent, accurate to 2 decimal places
    \end{itemize}
    \begin{enumerate}
        \item \( y = 127e^{-0.35t} \)
        \item \( y = 16e^{0.4t} \)
        \item \( y = 17250 e^{0.24t} \)
        \item \( y = 4700 e^{-0.07t} \)
    \end{enumerate}
\end{puzzle}





Identify if the function represents exponential growth, exponential decay, linear growth, or linear decay. In each case write the function and find the value at the indicated time.

\begin{puzzle}

    A house was purchased for \$350,000 in the year 2010. The value has been increasing by \$7,000 per year. Write the function and find the value of the house after 5 years.
\end{puzzle}

\begin{puzzle} A house was purchased for \$350,000 in the year 2010. The value has been increasing at the rate of 2\% per year. Write the function and find the value of the house after 5 years.
\end{puzzle}

\begin{puzzle} A lab purchases new equipment for \$50,000. Its value depreciates over time. The value decreases at the rate of 6\% annually. Write the function and find the value after 10 years.
\end{puzzle}

\begin{puzzle} A lab purchases new equipment for \$50,000. Its value depreciates over time. The value decreases by \$3000 annually. Write the function and find the value after 10 years.
\end{puzzle}

\begin{puzzle} A population of bats in a cave has 200 bats. The population is increasing by 10 bats annually. Write the function. How many bats live in the cave after 7 years?
\end{puzzle}

\begin{puzzle} A population of bats in a cave has 200 bats. The population is increasing at the rate of 5\% annually. Write the function. How many bats live in the cave after 7 years?
\end{puzzle}

\begin{puzzle} A population of a certain species of bird in a state park has 300 birds. The population is decreasing at the rate of 7\% per year. Write the function. How many birds are in the population after 6 years?
\end{puzzle}

\begin{puzzle} A population of a certain species of bird in a state park has 300 birds. The population decreases by 20 birds per year. Write the function. How many birds are in the population after 6 years?
\end{puzzle}

Write the exponential growth or decay function and find the value at the indicated time.

\begin{puzzle} A population of 400 microbes increases at the continuous growth rate of 26\% per day. Write the function and find the number of microbes in the population at the end of 7 days.
\end{puzzle}

\begin{puzzle} The price of a machine needed by a production factory is \$28,000. The business expects to replace the machine in 4 years. Due to inflation the price of the machine is increasing at the continuous rate of 3.5\% per year. Write the function and find the value of the machine 4 years from now.
\end{puzzle}

\begin{puzzle} A population of an endangered species consists of 4000 animals of that species. The population is decreasing at the continuous rate of 12\% per year. Write the function and find the size of the population at the end of 10 years.
\end{puzzle}

\begin{puzzle} A business buys a computer system for \$12000. The value of the system is depreciating and decreases at the continuous rate of 20\% per year. Write the function and find the value at the end of 3 years.
\end{puzzle}
