\section{Present Value Puzzles}

\begin{puzzle}
    Shawn has won a lottery paying him \$10,000 per month for the next 20 years. He'd rather have the whole amount in one lump sum today. If the current interest rate is 8.2\%, how much money can he hope to get?
    \end{puzzle}
    
    \begin{puzzle}
    Sonya bought a car for \$15,000. Find the monthly payment if the loan is to be amortized over 5 years at a rate of 10.1\%.
    \end{puzzle}
    
    \begin{puzzle}
    You determine that you can afford \$250 per month for a car. What is the maximum amount you can afford to pay for a car if the interest rate is 9\% and you want to repay the loan in 5 years?
    \end{puzzle}
    
    \begin{puzzle}
    Compute the monthly payment for a house loan of \$200,000 to be financed over 30 years at an interest rate of 10\%.
    \end{puzzle}
    
    \begin{puzzle}
    If the \$200,000 loan in the previous problem is financed over 15 years rather than 30 years at 10\%, what will the monthly payment be?
    \end{puzzle}
    
    \begin{puzzle}
    Friendly Auto offers Jennifer a car for \$2000 down and \$300 per month for 5 years. Jason wants to buy the same car but wants to pay cash. How much must Jason pay if the interest rate is 9.4\%?
    \end{puzzle}
    
    \begin{puzzle}
    The Gomez family bought a house for \$175,000. They paid 20\% down and amortized the rest at 11.2\% over a 30-year period. Find their monthly payment.
    \end{puzzle}
    
    \begin{puzzle}
        Mr. and Mrs. Wong purchased their new house for \$350,000. They made a down payment of 15\%, and amortized the rest over 30 years. If the interest rate is 9\%, find their monthly payment.
        \end{puzzle}
        
        \begin{puzzle}
        A firm needs a piece of machinery that has a useful life of 5 years. It has an option of leasing it for \$10,000 a year, or buying it for \$40,000 cash. If the interest rate is 10\%, which choice is better?
        \end{puzzle}
        
        \begin{puzzle}
        Jackie wants to buy a \$19,000 car, but she can afford to pay only \$300 per month for 5 years. If the interest rate is 6\%, how much does she need to put down?
        \end{puzzle}
        
        \begin{puzzle}
        Vijay’s tuition at Stanford for the next year is \$32,000. His parents have decided to pay the tuition by making nine monthly payments. If the interest rate is 9\%, what is the monthly payment?
        \end{puzzle}
        
        \begin{puzzle}
        Glen borrowed \$10,000 for his college education at 8\% compounded quarterly. Three years later, after graduating and finding a job, he decided to start paying off his loan. If the loan is amortized over five years at 9\%, find his monthly payment for the next five years.
        \end{puzzle}
        