\section{Chapter Review}

\begin{puzzle}
    Two dice are rolled. Find the probability that the sum of the dice is
    \begin{enumerate}
        \item four
        \item five
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    A jar contains 3 red, 4 white, and 5 blue marbles. If a marble is chosen at random, find the following probabilities:
    \begin{enumerate}
        \item \( P(\text{red or blue}) \)
        \item \( P(\text{not blue}) \)
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    A card is drawn from a standard deck. Find the following probabilities:
    \begin{enumerate}
        \item \( P(\text{a jack or a king}) \)
        \item \( P(\text{a jack or a spade}) \)
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    A basket contains 3 red and 2 yellow apples. Two apples are chosen at random. Find the following probabilities:
    \begin{enumerate}
        \item \( P(\text{one red, one yellow}) \)
        \item \( P(\text{at least one red}) \)
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    A basket contains 4 red, 3 white, and 3 blue marbles. Three marbles are chosen at random. Find the following probabilities:
    \begin{enumerate}
        \item \( P(\text{two red, one white}) \)
        \item \( P(\text{first red, second white, third blue}) \)
        \item \( P(\text{at least one red}) \)
        \item \( P(\text{none red}) \)
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    Four coins are tossed. Find the following probabilities:
    \begin{enumerate}
        \item \( P(\text{All heads}) \)
        \item \( P(\text{1 head and 3 tails}) \)
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    Consider three coins tossed. Find the following:
    \begin{enumerate}
        \item \( P(\text{both heads and tails appear } | \text{ first toss is a head}) \)
        \item \( P(\text{all tails } | \text{ both heads and tails appear}) \)
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    Mrs. Rossetti is flying from San Francisco to New York. On her way to the San Francisco Airport she encounters heavy traffic and determines that there is a 20\% chance that she will be late to the airport and will miss her flight. Even if she makes her flight, there is a 10\% chance that she will miss her connecting flight at Chicago. What is the probability that she will make it to New York as scheduled?
\end{puzzle}

\begin{puzzle}
    At a college, twenty percent of the students take history, thirty percent take math, and ten percent take both. What percent of the students take at least one of these two courses?
\end{puzzle}

\begin{puzzle}
    In a T-maze, a mouse may run to the right (R) or may run to the left (L). A mouse goes up the maze three times, and events E and F are described as follows:
    \begin{itemize}
        \item \( E \): Runs to the right on the first trial
        \item \( F \): Runs to the left two consecutive times
    \end{itemize}
    Determine whether the events E and F are independent.
\end{puzzle}

\begin{puzzle}
    A college has found that 20\% of its students take advanced math courses, 40\% take advanced English courses and 15\% take both advanced math and advanced English courses. If a student is selected at random, what is the probability that
    \begin{enumerate}
        \item They are taking English given that they are taking math?
        \item They are taking math or English?
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    If there are 35 students in a class, what is the probability that at least two have the same birthday?
\end{puzzle}

\begin{puzzle}
    A student feels that her probability of passing accounting is .62, of passing mathematics is .45, and her passing accounting or mathematics is .85. Find the probability that the student passes both accounting and math.
\end{puzzle}

\begin{puzzle}
    There are nine judges on the U. S. Supreme Court of which six are conservative and three liberal. This year the court will act on six major cases. What is the probability that out of six cases the court will favor the conservatives in at least four? Hint: all the information you need is here.
\end{puzzle}

\begin{puzzle}
    Five cards are drawn from a deck. Find the probability of obtaining
    \begin{enumerate}
        \item four cards of a single suit
        \item two cards of one suit, two of another suit, and one from the remaining
        \item a pair (e.g., two aces and three other cards)
        \item a straight flush (five in a row of a single suit but not a royal flush)
    \end{enumerate}
\end{puzzle}


\begin{puzzle}
    The following table shows a distribution of drink preferences amost the faculty and staff.

    \begin{center}
        \begin{tabular}{|l|c|c|c|c|}
            \hline
                        & Coke (C) & Pepsi (P) & Seven Up (U) & TOTALS \\
            \hline
            Staff (S)   & 60       & 50        & 22           & 132    \\
            Faculty (F) & 50       & 40        & 18           & 108    \\
            \hline
            TOTALS      & 110      & 90        & 40           & 240    \\
            \hline
        \end{tabular}
    \end{center}

    The events \( S, F, C, P \) and \( U \) are defined as Staff, Faculty, Coca Cola, Pepsi, and Seven Up, respectively. Find the following:
    \begin{enumerate}
        \item \( P(F | U) \)
        \item \( P(P | F) \)
        \item \( P(C | S) \)
        \item \( P(S | P \cup C) \)
        \item Are the events \( F \) and \( U \) mutually exclusive?
        \item Are the events \( F \) and \( U \) independent?
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    At a clothing outlet 20\% of the clothes are irregular, 10\% have at least a button missing and 4\% are both irregular and have a button missing. If Martha found a dress that has a button missing, what is the probability that it is irregular?
\end{puzzle}

\begin{puzzle}
    A pack of dogs consists of four Aussies, three Heelers and two Border Collies. Three dogs are chosen at random. Find the following probabilities:
    \begin{enumerate}
        \item \( P(\text{two Aussies and one Heeler}) \)
        \item \( P(\text{at least one Aussie}) \)
        \item \( P(\text{one of each}) \)
        \item\( P(\text{no Border Collies}) \)
    \end{enumerate}
\end{puzzle}

\begin{puzzle}
    A coin is tossed three times, and the events \( E \) and \( F \) are as follows.
    \[ E: \text{It shows a head on the first toss} \]
    \[ F: \text{Never turns up a tail} \]
    Are the events \( E \) and \( F \) independent?
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .6 \) and \( P(F) = .4 \) and \( E \) and \( F \) are mutually exclusive, find \( P(E \text{ and } F) \).
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .5 \) and \( P(F) = .3 \) and \( E \) and \( F \) are independent, find \( P(E \cup F) \).
\end{puzzle}

\begin{puzzle}
    If \( P(F') = .9 \) and \( P(E | F) = .36 \) and \( E \) and \( F \) are independent, find \( P(E) \).
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .4 \) and \( P(E \text{ or } F) = .9 \) and \( E \) and \( F \) are independent, find \( P(F) \).
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .4 \) and \( P(F | E) = .5 \), find \( P(E \text{ and } F) \).
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .6 \) and \( P(E \text{ and } F) = .3 \), find \( P(F | E) \).
\end{puzzle}

\begin{puzzle}
    If \( P(E) = .3 \) and \( P(F) = .4 \) and \( E \) and \( F \) are independent, find \( P(E | F) \).
\end{puzzle}
